---
title: "Research Stats 406"
author: "Muhammad Hafizudeen Mohamad Saman, hafizudn, Hannah Daane, hdaane"
date: "Due Oct. 3999 at midnight"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(3910349)
#install.packages("tidyverse")
#install.packages("ggplot2")
library(tidyverse)
```

```{r}
shootings <- as.data.frame(read.csv("school_shooting_db_20200316.csv"))
shootings$Summary <- NULL
shootings$Narrative..Detailed.Summary..Background. <- NULL
shootings$Sources <- NULL
shootings$Shooter.Name <- NULL
shootings$School <- NULL



# summary(shootings)
# shootings$City <- as.factor(shootings$City)
# shootings$State <- as.factor(shootings$State)
# shootings$Gender.of.Victims..M.F.Both. <- as.factor(shootings$Gender.of.Victims..M.F.Both.)
# shootings$Victim.s.Affiliation.w..School <- as.factor(shootings$Victim.s.Affiliation.w..School)
# shootings$Victims.Race <- as.factor(shootings$Victims.Race)
# shootings$Victim.Ethnicity <- as.factor(shootings$Victim.Ethnicity )
# shootings$Targeted.Specific.Victim.s. <- as.factor(shootings$Targeted.Specific.Victim.s.)
# shootings$Random.Victims <- as.factor(shootings$Random.Victims)
# shootings$Domestic.Violence..Y.N. <- as.factor(shootings$Domestic.Violence..Y.N.)
# shootings$Suicide..Shooter.was.only.victim..Y.N..N.A <- as.factor(shootings$Suicide..Shooter.was.only.victim..Y.N..N.A)


shootings <- as.data.frame(unclass(shootings),stringsAsFactors=TRUE)
summary(shootings)   
```
# Data

The data that we are going to use to analyze the suicide trend of school shooter over the year in this project is provided by the Center for Homeland Defense and Security (CHDS) at the Naval Postgraduate School. This data recorded the instances of school shootings that happen over the span of 50 years in the United States, ranging from 1970 to 2020. The data provides substantial information concerning each occasion of school shooting, including detailed location, time, victims, shooter, firearm used, hostages, and even a brief summary on what happened in each instance of shooting. 

The shooter information is available with several descriptors: Number of shooters, their names, type of attack (whether the shooting is accidental, gang-related, robbery, targeted attack, mental health related, or escalation of dispute), action immediately after shooting, whether the shooters have been bullied, their age, gender, race, ethnicity, affiliation with school, hostages taken, whether or not they have accomplice who did not fire a gun, and whether or not they attempted suicide after shooting. We are interested in looking deeper into the last variable (suicide attempts by the shooters) and analyze its trend over the five decades this data has been collected. We are also intended in looking at the shooters' ages to see how much each age group's suicide rate change over these years.

The suicide attempt by shooter column in the data consist of 5 different values: Y= Yes, N=No,Officer Involved = whether an officer is the shooter, Unknown = unknown, and N/A= no data is collected. There are 16 N/A's, 1 Unknown, and 24 Officer involved. In this paper, we are going to remove these three types of observation from our analysis and only look at the known values for the suicide attempts by non-officer shooters. Similarly, observations with no data on shooter's age will not be used in the age analysis. As for the trend increase, we will only look at the year as a whole. 1/1/2019 is equivalent to 12/31/2019, but is different from 1/1/2020.

Some data are not accurate. For example, there is a shooter who is listed as 40888 years old, which drastically shifted the mean of the shooter's age to 59.5 years old. These data are checked against the source and is replaced by its actual value, given that it is stated in the source. Also, the data for 2020 is not completed, as there are only observations for shooting in January. Since we are analyzing the data by the year, we have to exclude 2020 from our analysis.

A preliminary inspection of the data shows an increasing pattern in the suicide attempts by the shooter over the years, as given in the figure below. The number of suicide attempts is likely to be correlated with the number of shootings in that particular year. Therefore, to account for this, we will also look at the percentage of suicide attempts in addition to the raw number.
```{r}
#Change the value of the outlier
#naremoved <- shootings[!is.na(shootings$Shooter.Age),]
#naremoved$Shooter.Age[which(naremoved$Shooter.Age == 40888)] <- 12

library(ggplot2)
#Add Year Column
shootings$Year <- format(as.Date(shootings$Date, format="%m/%d/%Y"),"%Y")
shootings$Shooter.Age[which(shootings$Shooter.Age == 40888)] <- 12
ggplot(data = shootings) +
    geom_col(mapping = aes(x = Year, y = Suicide..or.attempted.suicide..by.Shooter..Y.N.)) + ylab("Number of Suicide Attempts")
```

```{r}
shootings$suicide <- ifelse((shootings$Suicide..or.attempted.suicide..by.Shooter..Y.N. == 'Y'), TRUE, FALSE)
str(shootings)
lm.rv <- lm(suicide~Killed..includes.shooter.+ Wounded + Total.Injured.Killed.Victims +
Victim.s.age.s. + Race + Shooter.Gender + Shooter.Age + Shooter.Ethnicity + Shooter.s.Affiliation.with.School, data = shootings)
lm.rv <- lm(suicide~Shooter.Ethnicity, data = shootings)
summary(lm.rv)
mean(shootings$Total.Injured.Killed.Victims)
```

#Hannah Part

```{r}
suicide.by.age <- shootings[,c("Shooter.Age", "suicide")]
suicide.by.age$count <- 1
agg.age_maybe  <- 
  mutate(suicide.by.age, Age.Shooter = Shooter.Age) %>% group_by(Shooter.Age, Age.Shooter) %>% 
      summarize_all(sum) %>% 
      select(-Shooter.Age)
#agg.age_maybe$count <- 1
agg.age_maybe$percentage <- agg.age_maybe$suicide/agg.age_maybe$count


ggplot(agg.age_maybe) +
  geom_col(mapping = aes(x = Age.Shooter, y = percentage)) + xlab("Shooter's Age")  + ylab("Percentage of Suicide")


summary(lm(formula = percentage~Age.Shooter, data = agg.age_maybe))
median(shootings$Shooter.Age, na.rm=TRUE)
```
New hannah code
```{r}
mean_data <-mean(agg.age_maybe$percentage)
var_data <- var(agg.age_maybe$percentage)
  (alpha = (((1 - mean_data) / var_data) - (1/mean_data)) * (mean_data^2))  
  (beta = alpha * ((1/mean_data) - 1))



f <- function() {rbeta(5, alpha, beta)}
 g <- function() { rnorm(5, 0 , 1)}
 
 estimate_rejection_prob <- function(f, g) {
  k <- 1000
  rejections <- replicate(k, {
    x <- f()
    y <- g()
    c(t.test(x, y)$p.value < 0.1,
      wilcox.test(x, y)$p.value < 0.1) 
  })
  
  est.t <- mean(rejections[1,])
  ci.t <- binom.test(sum(rejections[1,]), k, conf.level = 0.95)$conf.int
  
  est.w <- mean(rejections[2,])
  ci.w <- binom.test(sum(rejections[2,]), k, conf.level = 0.95)$conf.int
  
  matrix(c(est.t, ci.t, est.w, ci.w), nrow = 2, byrow = TRUE, 
         dimnames = list(test = c("t-test", "WMW"), value = c("Estimate", "Lower", "Upper")))
}
 
 #difference in power
 estimate_rejection_prob(f,g)

 #Boxplot comparing suicide and age
mutate(shootings, Age = Shooter.Age) %>%
ggplot() +
  geom_boxplot(mapping = aes(x = suicide, y = Age))


shootings %>%
ggplot() +
  geom_bar(mapping = aes(x = Year, fill = suicide)) +
  theme(axis.text.x = element_text(angle = 45, vjust = .7, hjust= .5))
#Median Year
median(shootings$Year)
#Mean Year
mean(shootings$Year)


#Percentage of Schoool Shootings with Suicides 
sum(shootings$suicide == TRUE) / sum(shootings$suicide == FALSE | TRUE)

```


#Hafiz Part



```{r}
#removing 2020 from the data
shootings <- shootings[1:1507,]

library(dplyr)
shootings$Year <- as.numeric(shootings$Year)
suicide.by.year <- as.data.frame(shootings)
suicide.by.year <- suicide.by.year[,c("Year", "suicide")]
suicide.by.year$count <- 1
agg.suicide <- suicide.by.year %>% mutate(year = floor(Year/1)*1) %>% 
      group_by(year) %>% 
      summarize_all(sum) %>% 
      select(-Year)
agg.suicide$percentage <- agg.suicide$suicide/agg.suicide$count
summary(lm(percentage ~ year, agg.suicide, weights = agg.suicide$count))

plot(agg.suicide$year,agg.suicide$percentage)

agg.suicide
```


# WLS model to fit linear regression on the number of suicide 
```{r}
lm.numberofsuicide.vs.year <- lm(suicide~year, agg.suicide, weights = agg.suicide$count)
summary(lm.numberofsuicide.vs.year)
```

# WLS model to fit linear regression on the PERCENTAGE of suicide 
```{r}
lm.percofsuicide.vs.year <- lm(percentage~year, agg.suicide, weights = agg.suicide$count)
summary(lm.percofsuicide.vs.year)

```
#Ordinary Least Squares for number of suicide vs year
```{r}
olm.numberofsuicide.vs.year <- lm(suicide~year, agg.suicide)
summary(olm.numberofsuicide.vs.year)
```


```{r}
##Simulation for WLS (Using real data)
#sample 20 years from 50yrs
#10000 times
beta_1_number <- c()
beta_1_percentage <- c()
for(i in 1:10000){
  sample_idx <- sample(1:nrow(agg.suicide), 20)
  sample_data <- agg.suicide[sample_idx,]
  lm.model <- lm(suicide~year, sample_data, weights = count)
  lm.model2 <- lm(percentage~year, sample_data, weights = count)
  beta_1_number[i] <- lm.model$coefficients[2]
  beta_1_percentage[i] <- lm.model2$coefficients[2]
}
hist(beta_1_number)
hist(beta_1_percentage)
quantile(beta_1_number, c(0.025,0.975))
quantile(beta_1_percentage, c(0.025,0.975))
sample(agg.suicide,1)

##Simulation for WLS (Using SIMULATED data)
#sample 20 years from 50yrs

ols.std.err <- c()
wls.std.err <- c()
wls.beta.1 <- c()
#creating simulated agg.suicide
for(i in 1:1000){
  n=rep(1:100,2)
  sigma2 = n^1.3
  eps = rnorm(n,mean=0,sd=sqrt(sigma2))
  y=a+b*n + eps
  datas <- data.frame(cbind(n,y))
  lm.model <- lm(y~n, datas)
  beta_1_number[i] <- lm.model$coefficients[2]
  ols.std.err[i] <- summary(lm.model)$coefficients[2,2]
  
  #creating weights
  wt <- 1/sigma2
  lm.model2 <- lm(y~n, datas, weights = wt)
  wls.beta.1[i] <- lm.model2$coefficients[2]
  wls.std.err[i] <- summary(lm.model2)$coefficients[2,2]
}

data2 <- data.frame(ols.std.err, wls.std.err)
#density(x,'OLS Std Error','density')
ggplot(data2) + 
  geom_density(aes(ols.std.err), col= "blue") + xlab('Standard Error') + ylab('density') + geom_density(aes(wls.std.err), col= "red")

```
```{r}
ggplot(agg.age_maybe, aes(x=Shooter.Age, y= percentage)) + geom_point() + xlab("Shooter's Age") + ylab("Percentage of Suicide") + geom_smooth(method='lm')


ggplot(agg.age_maybe, aes(x=Shooter.Age, y= suicide)) + geom_point() + xlab("Shooter's Age") + ylab("Number of Suicide") + geom_smooth(method='lm') 
```


#Suicide percentage between Young and Adult
```{r}
#data has to have age of shooter, suicide number for each age group, and count number for each age group
welch.t.test <- function(data){
  age.group <- data
  age.group$young <- ifelse(age.group$Age.Shooter <18, "Young", "Adult")
  age.group$young <- as.factor(age.group$young)
  young <- age.group[which(age.group$young == "Young"), ]
  adult <- age.group[which(age.group$young == "Adult"), ]
  t.test(young$percentage, adult$percentage)
}
ttest.age <- welch.t.test(agg.age_maybe)
ttest.age


#simulation`
#agg.age_maybe$young <- ifelse(agg.age_maybe$Age.Shooter <18, "Young", "Adult")
#agg.age_maybe$young <- as.factor(agg.age_maybe$young)
#age.group <- (agg.age_maybe%>% group_by(young) %>% summarize_all(sum))

#genderweight %>%
#  group_by(group) %>%
#  get_summary_stats(weight, type = "mean_sd")
```



We only have about 50 data points, so our regression might not be the most accurate. So we should use bootstrap to increase the number of observations. 

Question 1: Each year has different number of shootings. How to assign weights to years with more shooting?

Question 2: Should we use bootstrap to get more data? How to implement bootstrap if the data is not iid?
